---
title: "Process QC Output"
author: "Sander Wuyts"
date: "04 September 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(stringr)
library(tidyverse)

wd = "~/serverUA/sander/carrot_isolates/03_quality_control/"

setwd(wd)
quastReportFile = "out/transposed_report.tsv" 
ncbi_summary_file = "../02_download_ncbi_genomes/assembly_summary.txt"

# read quast report file
report = read.table(quastReportFile, header=T, sep="\t", comment.char="", quote="")
# column names: replace any number of points by one underscore
names(report) = str_replace_all(names(report), "\\.+", "_")
# column names: remove trailing underscore
names(report) = str_replace(names(report), "\\_$", "")
# column names: replace X by N (number of...)
names(report) = str_replace(names(report), "X", "N")

# read NCBI data file
ncbi_data = read.delim(ncbi_summary_file, sep = "\t", skip = 1) %>%
  rename(assembly_accession = `X..assembly_accession`)
# assembly accession numbers: take only part before the dot (to match accession numbers in quast table)
ncbi_data$assembly_accession = str_match(ncbi_data$assembly_accession, "(.*)( |\\.)")[,2]

# Read in checkM data
checkm <- read_tsv("out_checkm/results.tsv") %>%
  rename(Assembly = `Bin Id`)

# make super table with all info (ncbi and quast)
genomeTable = merge(report, ncbi_data, by.x="Assembly", by.y="assembly_accession", all.x=T) %>%
  left_join(checkm %>% select(-GC))

# make extra column with species name only (no strain names)
genomeTable <- genomeTable %>%
  mutate(species = str_match(organism_name, " ([a-z]*)")[,2])
# give species name "isolate" to the isolate genomes
genomeTable$species[is.na(genomeTable$species)] = "isolate"
# make extra column with genus name only
genomeTable <- genomeTable %>%
  mutate(strain = str_split_fixed(infraspecific_name, pattern = "=", n = 2)[,2]) %>%
  mutate(strain = if_else(strain == "", "isolate", strain)) %>%
  mutate(genus = str_match(organism_name, "([A-z]*) ")[,2])
genomeTable$genus[is.na(genomeTable$genus)] = "isolate"




genomeTable <- genomeTable %>% 
  mutate(species = if_else(str_detect(Assembly,"Mudanjiangensis"), "mudanjiangensis", 
                                         if_else(str_detect(Assembly, "Herbarum"), "herbarum", species))) %>%
  mutate(fromNCBI = if_else(species == "isolate", "isolate", 
                            if_else(species == "mudanjiangensis", "inhousRefgenome", "NCBIrefgenome")))

```

# Data distribution

FASTQ-filepairs: 70
Assembled genomes incl. L mudan: 66

```{r}
genomeTable %>%
  group_by(fromNCBI) %>%
  summarise(count = n())
```



```{r}
genomeTable %>%
  ggplot(aes(x = fromNCBI, fill = fromNCBI)) +
  geom_bar() 
```


# N50 and total genome length

Let's check the distribution of the N50 values:

```{r, echo=FALSE}
ggplot(data=genomeTable, aes(x=N50)) +
  geom_density(aes(group=NULL)) +
  geom_jitter(aes(y=(-0.05 - 0.1*ifelse(species=="isolate", 1, 0)),
                  col=fromNCBI, width=0, height=0.01),alpha=0.8) +
  scale_x_log10() +
  scale_color_brewer(palette="Paired", name="own isolate") +
  geom_vline(xintercept=2000)
```

There's a few isolates with a very short N50. Let's put a cutoff of N50  > 2000 bp to exclude these

Let's plot the quality parameters: N50 values and number of N's per 100.000 bases. 

```{r, echo=FALSE}
ggplot(data=genomeTable, aes(x=N_N_s_per_100_kbp, y=N50, col=fromNCBI)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  scale_color_brewer(palette="Paired", name="species") +
  geom_hline(yintercept=2000) +
  geom_vline(xintercept=1400)
```

Let's put a cutoff at around 1400.

```{r}
# Make extra variable with selected genomes
genomeTable$useGenome = genomeTable$N50>2000 & genomeTable$N_N_s_per_100_kbp<1400

# Summary
sprintf("total number of genomes: %i", nrow(report))
sprintf("number of genomes that pass QC: %i", sum(genomeTable$useGenome))
```

Now we can plot genome parameters: GC content and genome length: 

```{r, echo=FALSE}
ggplot(data=genomeTable, aes(x=Total_length, y=GC, col=fromNCBI)) +
  geom_point(alpha=0.8) +
  scale_color_brewer(palette="Paired", name="species") +
  geom_vline(xintercept=4e6)
```

There's one with a much bigger genome size than the rest.

```{r}
ggplot(data=genomeTable, aes(x=fromNCBI, y=Total_length, col=fromNCBI)) +
  geom_jitter(height=0, width=0.3) +
  scale_color_brewer(palette="Paired", name="species") + 
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1))
```

Let's also remove this one

```{r}
# Make extra variable with selected genomes
genomeTable$useGenome = genomeTable$useGenome & genomeTable$Total_length <4e+06

# Summary
sprintf("total number of genomes: %i", nrow(report))
sprintf("number of genomes that pass QC: %i", sum(genomeTable$useGenome))
```

# Completeness

I've also ran checkm to check for completeness!

```{r}
ggplot(data=genomeTable, aes(x=fromNCBI, y=Completeness, col=useGenome)) +
  geom_jitter(height=0, width=0.3) +
  scale_color_brewer(palette="Paired", name="species") + 
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1)) +
  geom_hline(yintercept = 95)
```

Many isolates show a low completeness. Let's remove all of them under 95% completeness!

```{r}
# Make extra variable with selected genomes
genomeTable$useGenome = genomeTable$useGenome & genomeTable$Completeness > 95

# Summary
sprintf("total number of genomes: %i", nrow(report))
sprintf("number of genomes that pass QC: %i", sum(genomeTable$useGenome))
```

```{r}
genomeTable[genomeTable$useGenome,] %>%
  group_by(fromNCBI) %>%
  summarise(count = n())
```


```{r}
genomeTable[genomeTable$useGenome,] %>%
  ggplot(aes(x = fromNCBI, fill = fromNCBI)) +
  geom_bar() 
```



# Some other standard plots

```{r}
ggplot(data=genomeTable[genomeTable$useGenome,], aes(x=fromNCBI, y=GC, col=fromNCBI)) +
  geom_jitter(height=0, width=0.3) +
  scale_color_brewer(palette="Paired", name="species") + 
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1))
```

```{r}
ggplot(data=genomeTable[genomeTable$useGenome,], aes(x=fromNCBI, y=N_contigs, col=fromNCBI)) +
  geom_jitter(height=0, width=0.3) +
  scale_color_brewer(palette="Paired", name="species") + 
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1))
```

```{r}
ggplot(data=genomeTable[genomeTable$useGenome,], aes(x=fromNCBI, y=Largest_contig, col=fromNCBI)) +
  geom_jitter(height=0, width=0.3) +
  scale_color_brewer(palette="Paired", name="species") + 
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1))
```

```{r}
ggplot(data=genomeTable[genomeTable$useGenome,], aes(x=fromNCBI, y=Total_length, col=fromNCBI)) +
  geom_jitter(height=0, width=0.3) +
  scale_color_brewer(palette="Paired", name="species") + 
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1))
```

```{r}
ggplot(data=genomeTable[genomeTable$useGenome,], aes(x=GC, y=Total_length, col=fromNCBI)) +
  geom_point() +
  scale_color_brewer(palette="Paired", name="species") 
```

# Statistics for paper

hier toevoegen optie in LGC?
```{r}
LGC_genera <- c("Lactobacillus",
                "Leuconostoc",
                "Weissella",
                "Pediococcus",
                "Fructobacillus",
                "Oenococcus",
                "isolate")

genomeTable <- genomeTable %>%
  mutate(inLGC = genus %in% LGC_genera)
```

Amount per species

```{r}
genomeTable %>%
  filter(useGenome == T) %>%
  filter(inLGC == T) %>%
  mutate(taxid = as.character(taxid)) %>%
  pull(taxid) %>%
  unique() %>%
  length()
```

# How many different fermentations?

```{r}
origins <- read_csv("labnames_and_origins.csv") %>%
  separate(Origin, into = c("Fermentation", "Day", "deleteme")) %>%
  select(-deleteme) %>%
  unite("Origin", Fermentation, Day, sep = " ") %>%
  mutate(PrevName = str_replace(PrevName, "/", "-"),
         PrevName = str_replace(PrevName, " ", "-"))
```

```{r}
genomeTable_origins <- genomeTable %>%
  filter(useGenome == T,
         fromNCBI == "isolate") %>%
  select(Assembly) %>%
  left_join(origins %>% rename(Assembly = PrevName)) %>%
  mutate(Labname = if_else(Assembly == "AMB-F197", "AMB-F197",
                           if_else(Assembly == "AMB-F249", "AMB-F249", Labname))) %>%
  mutate(Origin = if_else(Assembly == "AMB-F197", "FP02 D30",
                           if_else(Assembly == "AMB-F249", "FP27 D30", Origin)))
```

```{r}
genomeTable_origins %>%
  group_by(Origin) %>%
  summarise(numberOfFermentations = n()) %>%
  ggplot(aes(x = reorder(Origin, - numberOfFermentations), y = numberOfFermentations)) +
  geom_col() +
  coord_flip()
```


```{r}
genomeTable_origins %>%
  pull(Origin) %>%
  unique() %>%
  length()
```

29 Different fermentations.

```{r}
genomeTable <- genomeTable %>% 
  left_join(genomeTable_origins)
```


Make the genometable file!

```{r, echo=FALSE}
# Write file with all information of selected genomes
write.table(genomeTable[genomeTable$useGenome,], file="genomeTable_full.tsv",
            quote=F, col.names=T, row.names=F, sep="\t")

# Write table accession numbers and species names of selected genomes
write.table(genomeTable[genomeTable$useGenome,c("Assembly", "species")], file="genomeTable.tsv",
            quote=F, col.names=F, row.names=F, sep="\t")

# Write table with selected genomes only
write.table(genomeTable[genomeTable$useGenome,"Assembly"], file="genomesToUse.txt",
            quote=F, col.names=F, row.names=F, sep="\t")
```


```{r}
genomeTable %>%
  filter(useGenome == T,
         fromNCBI == "isolate") %>%
  ggplot(aes(y = Origin, x = GC)) +
  geom_jitter()
```

